\documentclass[12pt,a4paper]{book}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{float}

\setstretch{1.5}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue},
    commentstyle=\color{gray},
    stringstyle=\color{red},
    breaklines=true,
    showstringspaces=false,
    backgroundcolor=\color{lightgray!20},
    frame=single,
    rulecolor=\color{black}
}

\title{\textbf{\\[2cm]AI-Generated Media Detection Framework}\\
        \textbf{Complete Technical Documentation}\\[1cm]
        \normalsize{Multimodal Deep Learning for Detecting AI-Generated Images and Videos}}
\author{}
\date{\today}

\pagestyle{fancy}
\fancyhf{}
\rhead{AI Media Detection Framework}
\lhead{Technical Documentation}
\cfoot{\thepage}

\begin{document}

\maketitle

\begin{abstract}
This document provides a comprehensive technical analysis and detailed documentation of the AI-Generated Media Detection Framework. The system implements a sophisticated, cascading three-layer detection architecture for identifying AI-generated images and videos. The document covers every component, algorithm, threshold, model weight, data flow, and implementation detail found in the repository, organized for maximum clarity and technical understanding.
\end{abstract}

\tableofcontents
\newpage

% ===============================================
\chapter{Project Overview and Architecture}
\label{ch:overview}
% ===============================================

\section{System Purpose}
The AI-Generated Media Detection Framework is designed to analyze images and videos to determine whether they are authentic or generated by artificial intelligence systems. The system employs multiple complementary detection techniques and combines their results through advanced score fusion algorithms to achieve high accuracy while minimizing false positives and false negatives.

\section{Technology Stack}

\subsection{Backend}
\begin{itemize}
    \item Python 3.10 to 3.13 (tested versions)
    \item Flask (web framework for API server)
    \item PyTorch (deep learning framework)
    \item torchvision (computer vision utilities)
    \item Transformers library (Vision Transformer model)
    \item OpenCV (video processing and image manipulation)
    \item MediaPipe (human pose and hand detection)
    \item scikit-learn (machine learning utilities for meta-learner)
    \item C2PA library (content authenticity verification)
    \item Joblib (model serialization and deserialization)
    \item ONNX and onnx2pytorch (model format conversion)
    \item Numpy and Scipy (numerical computation)
\end{itemize}

\subsection{Frontend}
\begin{itemize}
    \item React 18.2.0 (user interface framework)
    \item Vite 5.1.0 (build tool and development server)
    \item React Router 6.22.0 (single-page application routing)
    \item TailwindCSS 3.4.1 (utility-first CSS framework)
    \item PostCSS (CSS preprocessing)
\end{itemize}

\subsection{Deployment and Server}
\begin{itemize}
    \item Waitress (production-grade WSGI application server)
    \item HuggingFace Hub (model repository and download)
\end{itemize}

\section{Project Directory Structure}
\label{sec:directory}

The repository is organized as follows:

\begin{lstlisting}[language=bash]
root/
├── backend/                              # Flask backend server
│   ├── app.py                            # Main Flask application
│   ├── combine_model.py                  # AI ensemble predictor
│   ├── custom_forensics.py               # Forensic analysis modules
│   ├── forensic.py                       # Report generation
│   ├── video_detect_standalone.py        # Video detection
│   ├── serve.py                          # Production server
│   ├── models/
│   │   ├── classifiers.py
│   │   ├── image.py
│   │   ├── rawnet.py
│   │   └── TMC.py
│   ├── src/
│   │   ├── c2pa_checker.py
│   │   └── synthid/
│   │       ├── service.py
│   │       ├── robust_extractor.py
│   │       └── __init__.py
│   ├── checkpoints/
│   │   ├── efficientnet.onnx
│   │   └── model.pth
│   ├── static/
│   ├── templates/
│   └── uploads/
├── frontend/                             # React frontend application
│   ├── src/
│   │   ├── App.jsx
│   │   ├── pages/
│   │   │   ├── Home.jsx
│   │   │   ├── Dashboard.jsx
│   │   │   ├── Report.jsx
│   │   │   ├── VideoDashboard.jsx
│   │   │   └── VideoReport.jsx
│   │   ├── index.css
│   │   └── main.jsx
│   ├── vite.config.js
│   ├── package.json
│   └── postcss.config.js
├── model_output/                        # Pre-trained model weights
│   ├── resnet50_finetuned_benchmark.pth
│   ├── ai_detector_meta_learner.joblib
│   ├── polynomial_transformer.joblib
│   └── synthid/
│       └── robust_codebook.pkl
├── scripts/windows/
│   ├── start-backend-dev.ps1
│   ├── start-backend-prod.ps1
│   ├── start-frontend-dev.ps1
│   └── health-check.ps1
├── requirements.txt
├── README.md
└── deployment_document.md
\end{lstlisting}

\section{Detection Architecture Overview}

The detection system employs a cascading three-layer architecture where each layer builds upon previous results:

\begin{itemize}
    \item \textbf{Layer 1 (C2PA)}: Checks for signed metadata indicating AI generation. If found, immediately returns AI Generated verdict.
    \item \textbf{Layer 2 (SynthID)}: Detects Google's SynthID watermarks embedded in AI-generated images. If detected with high confidence, returns AI Generated verdict.
    \item \textbf{Layer 3 (Deep Learning)}: Uses ensemble of neural networks combined with forensic analysis modules to make final determination.
\end{itemize}

This cascading approach ensures fast processing when high-confidence signals are present, while still providing sophisticated analysis for ambiguous cases.

% ===============================================
\chapter{Image Detection Pipeline}
\label{ch:image-detection}
% ===============================================

\section{Complete Image Analysis Flow}

When a user uploads an image for analysis, the system processes it through the following flow:

\begin{enumerate}
    \item File is received and validated (format, size)
    \item File is saved to temporary upload directory
    \item Layer 1: C2PA metadata check is performed
    \item If C2PA present: Return immediately with AI verdict
    \item Layer 2: SynthID watermark detection
    \item If watermark detected with high confidence: Return AI verdict
    \item Layer 3: Deep learning ensemble and forensics
    \item Combine scores from all available models and forensic modules
    \item Apply guards and thresholds
    \item Generate forensic report
    \item Return final verdict with confidence
\end{enumerate}

The complete flow is shown in Figure \ref{fig:imagearch}.

\section{Layer 1: C2PA Metadata Verification}
\label{sec:c2pa}

\subsection{Purpose and Location}
The C2PA (Content Authenticity and Provenance) checker verifies whether an image contains signed metadata asserting it was created by an AI system. C2PA is an industry standard for content authenticity developed by the Coalition for Content Provenance and Authenticity.

File location: \texttt{backend/src/c2pa\_checker.py}

\subsection{How C2PA Works}
C2PA metadata consists of digitally signed claims about how content was created. A manifest attached to the image contains:
\begin{itemize}
    \item Certificate issuer information
    \item Digital signatures
    \item Assertions about content creation method
    \item Timestamps
\end{itemize}

The key assertion for AI detection is the \texttt{digitalSourceType} which contains \texttt{trainedAlgorithmicMedia} for AI-generated content.

\subsection{Detection Logic}
The C2PA checker performs these steps:

\begin{enumerate}
    \item Attempts to read C2PA manifest from image
    \item If no manifest found: Returns \texttt{c2pa\_present = False}, continues to Layer 2
    \item If manifest found:
    \begin{itemize}
        \item Extracts issuer information
        \item Validates digital signatures
        \item Checks for AI-related assertions
        \item Returns \texttt{c2pa\_present = True}
    \end{itemize}
    \item Any metadata present triggers immediate AI verdict (no further analysis)
\end{enumerate}

\subsection{Response Format}
\begin{lstlisting}[language=Python]
{
    "c2pa_present": True or False,
    "available": True or False,
    "valid": True or False,
    "issuer": "certificate issuer" or None,
    "ai_generated": True or False,
    "raw_data": {...}  # Full manifest JSON
}
\end{lstlisting}

\subsection{Decision Policy}
If C2PA metadata is present, the system immediately returns verdict as ``AI Generated (C2PA Metadata Present)'' without proceeding to Layers 2 or 3. This is based on the principle that officially signed metadata is the strongest possible indicator of AI generation.

\section{Layer 2: SynthID Watermark Detection}
\label{sec:synthid}

\subsection{Overview}
SynthID is Google's watermarking technology for AI-generated images. Unlike traditional watermarks that are visually obvious, SynthID embeds information in a way that is imperceptible to humans but detectable through mathematical analysis.

Location: \texttt{backend/src/synthid/service.py} and \texttt{backend/src/synthid/robust\_extractor.py}

\subsection{SynthID Service Configuration}
The SynthID service contains calibration parameters that control detection sensitivity:

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{SynthID Calibration Parameters}
\begin{tabular}{|l|r|p{5cm}|}
\hline
\textbf{Parameter} & \textbf{Default} & \textbf{Meaning} \\
\hline
\texttt{SYNTHID\_MIN\_CONFIDENCE} & 0.50 & Minimum raw watermark detection confidence (0-1 scale) \\
\hline
\texttt{SYNTHID\_MIN\_PHASE\_MATCH} & 0.40 & Minimum phase correlation alignment score \\
\hline
\texttt{SYNTHID\_MIN\_CORRELATION\_MARGIN} & 0.0001 & Minimum margin above background threshold \\
\hline
\texttt{SYNTHID\_MIN\_CARRIER\_MATCH\_RATIO} & 0.15 & Minimum match strength for carrier frequency \\
\hline
\texttt{SYNTHID\_MIN\_SIGNAL\_SCORE} & 0.45 & Minimum composite signal quality score \\
\hline
\texttt{SYNTHID\_MIN\_STRUCTURE\_RATIO} & 0.78 & Image structure lower bound (normalized) \\
\hline
\texttt{SYNTHID\_MAX\_STRUCTURE\_RATIO} & 1.85 & Image structure upper bound (normalized) \\
\hline
\texttt{SYNTHID\_MAX\_MULTI\_SCALE\_STD} & 0.30 & Maximum acceptable multi-scale inconsistency \\
\hline
\end{tabular}
\end{table}

\subsection{Multi-Stage Extraction Pipeline}
\label{subsec:synthid-pipeline}

The SynthID extractor processes images through multiple sequential stages:

\subsubsection{Stage 1: Multi-Scale Analysis}
The image is analyzed at three different resolutions simultaneously:
\begin{itemize}
    \item 256 × 256 pixels (low resolution, fast)
    \item 512 × 512 pixels (medium resolution)
    \item 1024 × 1024 pixels (high resolution, detailed)
\end{itemize}

Processing the image at multiple scales allows detection of watermarks at various frequency bands. SynthID watermarks may manifest differently at different resolutions.

\subsubsection{Stage 2: Multi-Denoiser Fusion}
Before watermark extraction, the image is cleaned using four complementary denoising methods:

\begin{enumerate}
    \item \textbf{Wavelet Denoising}: Uses wavelets (db4, sym8, coif3) to decompose image into frequency bands and selectively reconstruct clean bands. Preserves edges and fine details.
    
    \item \textbf{Bilateral Filtering}: Edge-preserving smoothing that blurs flat regions while maintaining sharp boundaries. Used for smoothing without edge degradation.
    
    \item \textbf{Non-Local Means (NLM) Denoising}: Compares image patches globally to find similar regions and averages them. Highly effective but computationally expensive.
    
    \item \textbf{Wiener Filtering}: Spectral domain filtering that minimizes mean square error. Assumes Gaussian noise model.
\end{enumerate}

Results from all four denoising methods are combined (ensemble fusion) to produce a robust noise estimate.

\subsubsection{Stage 3: Known Carrier Frequency Detection}
The extractor maintains a database of 28 identified SynthID carrier frequencies extracted from analysis of 288 Gemini-generated images.

Known horizontal carriers (frequency pairs):
\begin{itemize}
    \item (48, 0), (96, 0), (192, 0), (210, 0), (238, 0)
\end{itemize}

Known vertical carriers:
\begin{itemize}
    \item (0, 88), (0, 176), (0, 192)
\end{itemize}

Diagonal combinations include carriers like:
\begin{itemize}
    \item (48, 88), (96, 88), (96, 176), and others
\end{itemize}

The system searches the image's frequency spectrum for these known carriers. If found, it indicates potential SynthID watermark.

\subsubsection{Stage 4: ICA/PCA-Based Watermark Separation}
Independent Component Analysis (ICA) or Principal Component Analysis (PCA) is applied to decompose the image into independent components. This technique attempts to separate the watermark signal from the natural image signal through blind source separation.

\subsubsection{Stage 5: Adaptive Thresholding}
The threshold for watermark detection is not fixed but adapts based on image content:
\begin{itemize}
    \item Image's background power spectrum is estimated
    \item Threshold is set relative to background noise level
    \item High noise images require stronger signals for detection
    \item Clean images allow lower signal thresholds
\end{itemize}

\subsection{Calibration Logic: Two Detection Paths}

The SynthID calibration logic uses two different sets of thresholds depending on whether the raw extractor detected a watermark:

\subsubsection{Path 1: Raw Detection Fired (Relaxed Requirements)}
If the underlying watermark extractor's raw detector triggers (internal binary classification), the system applies loosened thresholds:

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{SynthID Thresholds: Raw Detection Fired Path}
\begin{tabular}{|l|r|r|}
\hline
\textbf{Parameter} & \textbf{Strict} & \textbf{Relaxed (if raw fired)} \\
\hline
\texttt{min\_confidence} & ≥ 0.50 & ≥ 0.475 (95\% of 0.50) \\
\hline
\texttt{min\_phase\_match} & ≥ 0.40 & ≥ 0.368 (92\% of 0.40) \\
\hline
\texttt{min\_signal\_score} & ≥ 0.45 & ≥ 0.405 (90\% of 0.45) \\
\hline
\texttt{min\_correlation\_margin} & ≥ 0.0001 & ≥ 0.00007 (70\% of margin) \\
\hline
\texttt{min\_carrier\_match} & ≥ 0.15 & ≥ 0.1275 (85\% of 0.15) \\
\hline
\end{tabular}
\end{table}

This relaxed path prevents suppression of true positives due to environmental drift (differences in image characteristics across different software/hardware).

\subsubsection{Path 2: Normal Path (Strict Requirements)}
If raw detector did not fire, all parameters must pass strict thresholds:
\begin{itemize}
    \item \texttt{signal\_score} must be ≥ 0.45
    \item \texttt{raw\_confidence} must be ≥ 0.50 (expressed as 0-100 scale)
    \item \texttt{phase\_match} must be ≥ 0.40
    \item \texttt{correlation\_margin} must be ≥ 0.0001
    \item \texttt{carrier\_match\_ratio} must be ≥ 0.15
    \item \texttt{multi\_scale\_consistency} must be ≤ 0.30
    \item \texttt{structure\_ratio} must be within range [0.78, 1.85]
\end{itemize}

All conditions must pass simultaneously for watermark detection.

\subsection{Signal Score Composition}
The signal score is a composite measure combining multiple quality metrics:

\begin{equation}
\text{signal\_score} = \text{weighted combination of:}
\end{equation}

\begin{itemize}
    \item Raw confidence (normalized from 0.50-0.95 range to 0-1 scale)
    \item Phase match quality (0-1 scale)
    \item Correlation margin (0-1 scale, normalized)
    \item Structure ratio validity (penalty if outside normal range)
    \item Multi-scale consistency (penalty for high variation)
    \item Carrier match ratio strength (0-1 scale)
\end{itemize}

The exact weighting is determined by the robust extractor implementation, prioritizing confidence and phase match as primary indicators.

\subsection{Detection Output}
The SynthID analysis returns comprehensive information about the watermark detection:

\begin{lstlisting}[language=Python]
{
    "status": "complete",  # complete, error, unavailable
    "available": True,
    "is_watermarked": True or False,  # Final decision
    "is_watermarked_raw": True or False,  # Raw detector
    "message": "string description",
    "confidence": 0-100,  # Confidence as percentage
    "raw_confidence": 0-1,  # Raw confidence 0-1 scale
    "correlation": float,  # Cross-correlation peak
    "phase_match": float,  # Phase alignment (0-1)
    "structure_ratio": float,  # Content structure measure
    "carrier_strength": float,  # Watermark carrier strength
    "carrier_match_ratio": float,  # (0-1) match quality
    "multi_scale_consistency": float,  # (0-1) consistency
    "correlation_margin": float,  # Margin above baseline
    "signal_score": float,  # Composite quality (0-1)
    "signal_band": "strong" or "weak" or "negative",
    "threshold": float,  # Decision threshold used
    "calibration": {
        "applied": True,
        "min_confidence": 0.50,
        "min_phase_match": 0.40,
        "min_correlation_margin": 0.0001,
        "min_carrier_match_ratio": 0.15,
        "min_signal_score": 0.45,
        "structure_ratio_range": [0.78, 1.85],
        "max_multi_scale_std": 0.30
    },
    "source": "reverse-SynthID robust extractor"
}
\end{lstlisting}

\subsection{Signal Bands}
The detected watermark is classified into three signal bands:

\begin{itemize}
    \item \textbf{Strong}: Watermark detected with calibrated high confidence
    \item \textbf{Weak}: Watermark-like signal present but below calibrated threshold
    \item \textbf{Negative}: No significant watermark signal detected
\end{itemize}

\section{Layer 3: Deep Learning Ensemble and Forensic Analysis}
\label{sec:layer3}

\subsection{Overview}
Layer 3 combines AI neural network models with forensic analysis to make the final determination about image authenticity. This layer is only reached if C2PA and SynthID checks do not provide definitive answers.

File locations:
\begin{itemize}
    \item \texttt{backend/combine\_model.py} (main ensemble logic)
    \item \texttt{backend/custom\_forensics.py} (forensic modules)
\end{itemize}

\subsection{AI Model Ensemble}
\label{subsec:ensemble}

Three AI models work together, each contributing to the final probability score:

\subsubsection{ResNet50 Model}
\begin{itemize}
    \item Architecture: ResNet-50 with 50 layers, modified FC (fully connected) layer for binary classification
    \item Input size: Exactly 224 × 224 pixels
    \item Preprocessing: Standardization with ImageNet statistics
    \begin{itemize}
        \item Normalize means: [0.485, 0.456, 0.406]
        \item Normalize standard deviations: [0.229, 0.224, 0.225]
    \end{itemize}
    \item Output: Two values from FC layer (real class, AI class)
    \item Output index for AI: Configured via \texttt{RESNET\_AI\_INDEX} (default: 1)
    \item Weights file: \texttt{model\_output/resnet50\_finetuned\_benchmark.pth}
    \item Size: Approximately 100 MB
    \item Test-Time Augmentation: Optional horizontal flip for robustness
    \begin{itemize}
        \item Controlled by environment variable \texttt{ENABLE\_TTA} (default: enabled)
        \item Image is flipped horizontally, re-inferred, and averaged with original
        \item Improves accuracy at cost of 2x inference time
    \end{itemize}
    \item Status: Always available and used as baseline
\end{itemize}

\subsubsection{Vision Transformer (ViT) Model}
\begin{itemize}
    \item Pre-trained model: \texttt{dima806/ai\_vs\_real\_image\_detection}
    \item Source: Downloaded from HuggingFace model hub
    \item Access: Requires internet connection on first use (then cached locally)
    \item Output: Two values or sigmoid output for binary classification
    \item Output index for AI: Auto-detected or specified via \texttt{VIT\_AI\_INDEX}
    \item Size: Approximately 350 MB
    \item Disabling: Can be skipped via \texttt{SKIP\_VIT=1} environment variable
    \item Status: Optional; system works if unavailable but with reduced accuracy
    \item Advantage: Transformer architecture captures long-range dependencies in images
\end{itemize}

\subsubsection{Meta-Learner (Fusion Model)}
\begin{itemize}
    \item Type: Gradient boosting classifier or similar ensemble method
    \item Input: Features extracted from [ResNet score, ViT score]
    \item Processing: Polynomial feature transformation before feeding to learner
    \item Weights file: \texttt{model\_output/ai\_detector\_meta\_learner.joblib}
    \item Transformer file: \texttt{model\_output/polynomial\_transformer.joblib}
    \item Output: Calibrated AI probability (0-1 scale)
    \item Status: Optional; system uses simple average if unavailable
    \item Training: Meta-learner was trained on validation set to combine model outputs optimally
\end{itemize}

\subsection{Score Combination Logic}

The three models produce scores that must be intelligently combined. The logic adapts based on model availability:

\subsubsection{Case 1: Only ResNet Available}
\begin{lstlisting}[language=Python]
final_ai_score = resnet_score
# Legacy baseline behavior
\end{lstlisting}

\subsubsection{Case 2: ResNet + ViT Available, No Meta-Learner}
\begin{lstlisting}[language=Python]
base_blend = (resnet_score + vit_score) / 2.0
final_ai_score = base_blend
# Simple averaging when meta-learner unavailable
\end{lstlisting}

\subsubsection{Case 3: All Three Models Available}
\begin{lstlisting}[language=Python]
base_blend = (resnet_score + vit_score) / 2.0

# Create polynomial features
poly_features = polynomial_transformer.transform(
    [[resnet_score, vit_score]]
)

# Get meta-learner prediction
meta_prob = meta_learner.predict_proba(poly_features)[0, 1]

# Check for version mismatch (retraining issue)
if version_mismatch_detected:
    # Partial blend if versions don't match
    final_ai_score = 0.55 * base_blend + 0.45 * meta_prob
else:
    # Normal operation
    drift = abs(meta_prob - base_blend)
    
    if drift > 0.35:
        # Strong disagreement between base and meta
        final_ai_score = 0.50 * base_blend + 0.50 * meta_prob
    elif drift > 0.25:
        # Moderate disagreement
        final_ai_score = 0.40 * base_blend + 0.60 * meta_prob
    else:
        # Good agreement, trust meta-learner
        final_ai_score = meta_prob
\end{lstlisting}

The logic prioritizes the meta-learner when models agree closely, but reverts to averaging when there is disagreement to avoid overfitting to a single model.

\subsection{Forensic Analysis Modules}
\label{subsec:forensics}

Three independent forensic modules analyze specific types of AI artifacts:

\subsubsection{Module 1: Frequency Analysis}

\textbf{Purpose}: Detect checkerboard patterns and grid artifacts from bilinear upsampling used in image generation.

\textbf{Algorithm}:

\begin{enumerate}
    \item Convert image to grayscale
    \item Apply Hanning window function for FFT prep
    \item Compute 2D Fast Fourier Transform (FFT)
    \item Shift zero-frequency component to center
    \item Calculate power spectral density (PSD) in log scale
    \item Focus analysis on annular region (radius 22-88\% of image)
    \item Search for symmetric peak pairs in high-frequency region
    \item Measure grid alignment energy
    \item Calculate baseline noise threshold
\end{enumerate}

\textbf{Detection Criteria}:

The module looks for evidence of artificial upsampling:

\begin{itemize}
    \item Symmetric peak pairs at regular frequency intervals
    \item Grid-aligned energy concentrations (harmonic structure)
    \item Peak prominence above background
    \item Consistency of grid energy ratio
\end{itemize}

\textbf{Scoring Formula}:

\begin{equation}
\text{peak\_score} = \text{clip}\left(\frac{\text{prominence} - 0.90}{2.6}\right) \quad 40\% \text{ weight}
\end{equation}

\begin{equation}
\text{pair\_score} = \text{clip}\left(\frac{\text{pair\_count}}{6.0}\right) \quad 25\% \text{ weight}
\end{equation}

\begin{equation}
\text{grid\_score} = \text{clip}\left(\frac{\text{energy\_ratio} - 1.02}{0.45}\right) \quad 25\% \text{ weight}
\end{equation}

\begin{equation}
\text{alignment\_score} = \text{clip(axis\_alignment)} \quad 10\% \text{ weight}
\end{equation}

\begin{equation}
\text{ai\_score} = 0.40 \cdot \text{peak} + 0.25 \cdot \text{pair} + 0.25 \cdot \text{grid} + 0.10 \cdot \text{alignment}
\end{equation}

\textbf{Detection Trigger}:

AI is detected if:

\begin{equation}
\text{ai\_score} \geq 0.55 \text{ OR } (\text{pair\_count} \geq 4 \text{ AND } \text{energy\_ratio} > 1.18)
\end{equation}

\textbf{Output}:
\begin{lstlisting}[language=Python]
{
    "available": True,
    "ai_score": float,  # 0-1 scale
    "detected": True or False,
    "message": "string",
    "symmetric_peak_pairs": int,
    "grid_energy_ratio": float,
    "peak_prominence": float
}
\end{lstlisting}

\subsubsection{Module 2: Anatomy Analysis}

\textbf{Purpose}: Detect biomechanical impossibilities in human body and hand poses.

\textbf{Detection Method}: Uses MediaPipe to extract 2D landmarks for:
\begin{itemize}
    \item Body: 12 core joints (shoulders, elbows, wrists, hips, knees, ankles)
    \item Hands: 21 landmarks per hand (up to 2 hands detected)
\end{itemize}

\textbf{Anomaly Detection}:

\textbf{Body Anomalies}:

\begin{itemize}
    \item Limb length ratios
    \begin{itemize}
        \item Upper arm to lower arm ratio should be 1.0-1.35x
        \item Anomaly detected if ratio > 1.55x
        \item Hard limit: 2.10x (extremely unnatural)
        \item Upper leg to lower leg: Similar bounds
    \end{itemize}
    
    \item Joint angles
    \begin{itemize}
        \item Elbows and knees: Should be 8-178 degrees
        \item Acute angles < 8°: Unnatural, scored by formula: $\text{score} \times \frac{12 - \text{angle}}{12}$
        \item Straight angles 170-178°: Normal
        \item Hyper-extended > 178°: Suspicious, penalized
    \end{itemize}
    
    \item Frame outliers
    \begin{itemize}
        \item Landmarks should be in range [-0.2, 1.2] (normalized coordinates)
        \item Outside this range: Indicates detection error or body part off-frame
    \end{itemize}
\end{itemize}

\textbf{Hand Anomalies}:

\begin{itemize}
    \item Finger segment ratios
    \begin{itemize}
        \item Each finger has phalanges (bone segments)
        \item Ratios between segments: Normal 1.0-1.95x
        \item Hard limit: 3.20x
    \end{itemize}
    
    \item Palm-relative proportions
    \begin{itemize}
        \item Hand width to finger length: 0.45-2.40x range
        \item Outside this: Unnatural hand shape
    \end{itemize}
\end{itemize}

\textbf{Scoring}:

\begin{equation}
\text{body\_anomaly\_score} = 0.55 \cdot \text{ratio\_anom} + 0.30 \cdot \text{angle\_anom} + 0.15 \cdot \text{outlier\_anom}
\end{equation}

\textbf{Overall}:

\begin{equation}
\text{ai\_score} = 0.70 \cdot \text{body\_anom} + 0.30 \cdot \text{hand\_anom}
\end{equation}

\textbf{Output}:
\begin{lstlisting}[language=Python]
{
    "available": True or False,
    "ai_score": float,  # 0-1 scale
    "detected": True or False,
    "message": "string",
    "details": {
        "body": {...},
        "hands": {...},
        "messages": ["message1", "message2"]
    }
}
\end{lstlisting}

\subsubsection{Module 3: Entropy Analysis}

\textbf{Purpose}: Detect unnatural texture and detail distribution patterns characteristic of AI generation.

\textbf{Algorithm}:

\begin{enumerate}
    \item Resize image to maximum 512×512 pixels (maintains aspect ratio)
    \item Divide image into grid:
    \begin{itemize}
        \item 3×3 grid for small images (< 128×128)
        \item 4×4 grid for larger images
    \end{itemize}
    \item For each patch in grid:
    \begin{itemize}
        \item Compute Shannon entropy: $H = -\sum p_i \log(p_i)$ where $p_i$ are pixel intensity probabilities
        \item High entropy: Rich texture and detail
        \item Low entropy: Smooth, bland region
    \end{itemize}
    \item Calculate global statistics
    \begin{itemize}
        \item Center patches (3-9) entropy
        \item Edge patches entropy
        \item Center-edge gap
        \item Local entropy spread (variance)
        \item Fraction of low-entropy patches
    \end{itemize}
\end{enumerate}

\textbf{Detection Criteria}:

AI generators often produce unnatural entropy distributions:

\begin{itemize}
    \item Center regions have more detail than edges (``mushy background'')
    \item High local variation (inconsistent texture)
    \item Large bland background regions
\end{itemize}

\textbf{Scoring Formula}:

\begin{equation}
\text{center\_edge\_gap} = \text{mean}(\text{center\_entropy}) - \text{mean}(\text{edge\_entropy})
\end{equation}

\begin{equation}
\text{mushy\_bg\_score} = \text{clip}\left(\frac{\text{center\_edge\_gap} - 0.60}{2.4}\right) \quad 45\% \text{ weight}
\end{equation}

\begin{equation}
\text{spread\_score} = \text{clip}\left(\frac{\text{entropy\_spread} - 1.1}{2.4}\right) \quad 35\% \text{ weight}
\end{equation}

\begin{equation}
\text{bland\_bg\_score} = \text{clip}\left(\frac{\text{low\_entropy\_frac} - 0.35}{0.45}\right) \quad 20\% \text{ weight}
\end{equation}

\begin{equation}
\text{ai\_score} = 0.45 \cdot \text{mushy} + 0.35 \cdot \text{spread} + 0.20 \cdot \text{bland}
\end{equation}

\textbf{Detection Trigger}:

\begin{equation}
\text{detected} = (\text{ai\_score} \geq 0.45)
\end{equation}

\textbf{Guard Against Natural Low-Detail Images}:

If an image is naturally low-detail (e.g., architectural photo with simple background):

\begin{equation}
\text{if entropy\_spread} < 0.45 \text{ AND center\_gap} < 0.15:
\end{equation}

Then \texttt{bland\_score = 0} (remove bland penalty).

\textbf{Output}:
\begin{lstlisting}[language=Python]
{
    "available": True,
    "ai_score": float,  # 0-1 scale
    "detected": True or False,
    "message": "string",
    "global_entropy": float,
    "center_entropy": float,
    "edge_entropy": float,
    "local_entropy_spread": float,
    "low_entropy_fraction": float
}
\end{lstlisting}

\subsection{Forensic Score Calibration}
\label{subsec:forensic-calib}

The three forensic modules' scores are combined with the AI model score using weighted averaging:

\begin{equation}
\text{final\_forensic\_score} = \text{max}(\text{frequency}, \text{anatomy}, \text{entropy})
\end{equation}

The weighted combination:

\begin{lstlisting}[language=Python]
CUSTOM_WEIGHTS = {
    "resnet": 0.50,        # 50% from ResNet50
    "frequency": 0.20,     # 20% from frequency analysis
    "anatomy": 0.20,       # 20% from anatomy analysis
    "entropy": 0.10,       # 10% from entropy analysis
}
\end{lstlisting}

\textbf{Support Level Classification}:

\begin{itemize}
    \item \textbf{Supporting evidence}: Module score $\geq 0.60$ (strong signal)
    \item \textbf{Neutral}: $0.15 < $ score $< 0.60$ (ambiguous)
    \item \textbf{Negative evidence}: Module score $\leq 0.15$ (contradicts AI)
\end{itemize}

\textbf{Calibration Rules}:

\textbf{Rule 1: Multiple Forensic Support (2+ modules with score ≥ 0.60)}

\begin{lstlisting}[language=Python]
weighted_avg = (
    0.50 * resnet +
    0.20 * frequency +
    0.20 * anatomy +
    0.10 * entropy
)
forensic_avg = (frequency + anatomy + entropy) / 3.0
final_ai_prob = max(weighted_avg, 0.80*resnet + 0.20*forensic_avg)
\end{lstlisting}

When multiple forensic modules support AI detection, increase confidence in AI verdict.

\textbf{Rule 2: Single Forensic Support}

\begin{lstlisting}[language=Python]
if only_frequency_support:
    # Frequency can be compression artifacts
    # Be conservative with weighting
    final_ai_prob = 0.70*resnet + 0.30*weighted_avg
else:
    # Anatomy/Entropy support more reliable
    final_ai_prob = 0.85*resnet + 0.15*forensic_avg
\end{lstlisting}

Frequency artifacts can occur with compression, so weight it lower than anatomy/entropy.

\textbf{Rule 3: No Forensic Support (all negative or unavailable)}

\begin{lstlisting}[language=Python]
weighted_avg = (
    0.50 * resnet +
    0.20 * frequency +
    0.20 * anatomy +
    0.10 * entropy
)
final_ai_prob = 0.60*resnet + 0.40*weighted_avg

if max_module_score <= 0.20:
    final_ai_prob -= 0.10  # Reduce by 10%
elif forensic_avg <= 0.12:
    final_ai_prob -= 0.05  # Reduce by 5%
\end{lstlisting}

When forensic modules don't support AI verdict, lower the confidence in AI classification.

\subsection{Score Consensus and Guards}

After computing AI probability from models and forensics, additional guards check for model disagreement:

\subsubsection{Score Consensus Calculation}

\begin{equation}
\text{ai\_consensus} = \begin{cases}
0.80 \cdot \text{legacy\_ensemble} + 0.20 \cdot \text{resnet} & \text{if ViT unavailable} \\
0.50 \cdot \text{vit} + 0.30 \cdot \text{legacy\_ensemble} + 0.20 \cdot \text{resnet} & \text{if ViT available}
\end{cases}
\end{equation}

\begin{equation}
\text{real\_consensus} = 1.0 - \text{ai\_consensus}
\end{equation}

\begin{equation}
\text{score\_margin} = \text{ai\_consensus} - \text{real\_consensus}
\end{equation}

\subsubsection{Guard 1: ViT Disagreement Guard}

\textbf{Activation Conditions}:
\begin{itemize}
    \item ViT is highly confident AI ($\text{vit\_prob} \geq 0.86$)
    \item ResNet is confident real ($\text{resnet\_prob} \leq 0.22$)
    \item Strong disagreement: $\text{vit} - \text{resnet} \geq 0.55$
    \item At least mild forensic support exists (anatomy or entropy > 0.35)
\end{itemize}

\textbf{Correction}:

\begin{equation}
\text{guard\_ai\_prob} = 0.62 \cdot \text{vit} + 0.23 \cdot \text{legacy} + 0.15 \cdot \max(\text{forensic}, \text{freq})
\end{equation}

If \texttt{guard\_ai\_prob} > original \texttt{final\_ai\_prob}, use the guarded value.

\textbf{Purpose}: ViT may be right despite ResNet disagreement if forensic support exists. This guard boosts confidence in ViT's verdict.

\subsubsection{Guard 2: Real Disagreement Guard}

\textbf{Activation Conditions}:
\begin{itemize}
    \item ViT moderately confident AI ($\text{vit\_prob} \geq 0.66$)
    \item ResNet strongly confident real ($\text{resnet\_prob} \leq 0.18$)
    \item Disagreement: $\text{vit} - \text{resnet} \geq 0.42$
    \item Very weak forensic support (all modules < 0.22)
\end{itemize}

\textbf{Correction}:

\begin{equation}
\text{guard\_real\_prob} = 0.58 \cdot \text{resnet} + 0.22 \cdot \text{legacy} + 0.20 \cdot \text{forensic}
\end{equation}

\textbf{Purpose}: ResNet's confidence in real combined with weak forensics suggests false positive in ViT. Lower AI probability.

\subsection{Dynamic Decision Threshold}
\label{subsec:dynamic-threshold}

Unlike fixed thresholds, the decision threshold adapts based on:

1. Amount of forensic support
2. Presence of negative evidence
3. Model disagreement
4. Feature confidence margins
5. Availability of certain modules

\textbf{Base Threshold Selection}:

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Base Decision Thresholds by Forensic Support}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Forensic Support} & \textbf{With ViT} & \textbf{ResNet Only} \\
\hline
2+ modules (strong) & 0.54 & 0.57 \\
\hline
1 module (strong) & 0.58 & 0.62 \\
\hline
1 module (weak) & 0.62 & 0.66 \\
\hline
No support, 0 negative & 0.64 & 0.68 \\
\hline
No support, 1 negative & 0.70-0.72 & 0.74-0.76 \\
\hline
No support, 2+ negative & 0.76-0.80 & 0.80+ \\
\hline
\end{tabular}
\end{table}

\textbf{Dynamic Adjustments}:

\begin{enumerate}
    \item \textbf{High Model Disagreement}: If ResNet/ViT disagreement $\geq 0.55$ AND weak forensic support
    \begin{equation}
    \text{threshold} \mathrel{+}= 0.18
    \end{equation}
    
    \item \textbf{Anatomy Unavailable}: If anatomy module unavailable + weak forensics + 1+ negatives
    \begin{equation}
    \text{threshold} = 0.86
    \end{equation}
    
    \item \textbf{Strong ViT Confidence with Margin}: If ViT $\geq 0.85$ AND score\_margin $\geq 0.32$
    \begin{itemize}
        \item With forensic support: $\text{threshold} = \min(\text{threshold}, 0.66)$ (lower for confidence)
        \item Without support: $\text{threshold} = \max(\text{threshold}, 0.74)$ (raise to be safe)
    \end{itemize}
\end{enumerate}

\subsection{Final Confidence Calculation}

Once AI probability and threshold are determined, convert to 0-100 confidence scale:

\textbf{If AI Detected}:
\begin{equation}
\text{confidence} = 50 + 50 \cdot \frac{\text{ai\_prob} - \text{threshold}}{1.0 - \text{threshold}}
\end{equation}

\textbf{If Real Image}:
\begin{equation}
\text{confidence} = 50 + 50 \cdot \frac{\text{threshold} - \text{ai\_prob}}{\text{threshold}}
\end{equation}

This scaling ensures:
\begin{itemize}
    \item Score at threshold = 50\% confidence
    \item Score at 1.0 = 100\% confidence (for AI)
    \item Score at 0.0 = 100\% confidence (for Real)
\end{itemize}

\section{Layer 3 Decision Logic}

After computing all scores, guards, and threshold, final decision:

\begin{lstlisting}[language=Python]
if ai_probability >= final_threshold:
    verdict = "AI Image"
    confidence = calculate_confidence_ai(...)
else:
    verdict = "Real Image"
    confidence = calculate_confidence_real(...)
\end{lstlisting}

% ===============================================
\chapter{Video Detection Pipeline}
\label{ch:video-detection}
% ===============================================

\section{Overview}

Video detection is implemented as a frame-based analysis system. Rather than using optical flow or temporal consistency, the current implementation extracts frames and analyzes them individually, then aggregates results.

File location: \texttt{backend/video\_detect\_standalone.py}

\section{Video Processing Flow}

\subsection{Frame Sampling}

The video is processed as follows:

\begin{enumerate}
    \item Open video file using OpenCV's VideoCapture
    \item Determine total frame count
    \item Calculate sample indices: uniformly distribute 3 frames across the video
    \begin{itemize}
        \item Frame 1: At 0\% position (beginning)
        \item Frame 2: At 50\% position (middle)
        \item Frame 3: At 100\% position (end)
    \end{itemize}
    \item Iterate through all frames:
    \begin{itemize}
        \item Skip frames not in sample set (fast grabbing)
        \item Retrieve and process only sampled frames
    \end{itemize}
\end{enumerate}

\subsection{Per-Frame Processing}

For each sampled frame:

\begin{enumerate}
    \item Read frame from video (BGR format from OpenCV)
    \item Convert color space: BGR $\rightarrow$ RGB
    \item Resize to exactly 256×256 pixels
    \item Normalize pixel values by dividing by 255.0 (scale to [0, 1])
    \item Convert to PyTorch tensor with batch dimension
    \item Feed through EfficientNet encoder
\end{enumerate}

\subsection{EfficientNet Encoder}

The video encoder is based on EfficientNet architecture:

\begin{itemize}
    \item Format: ONNX (Open Neural Network Exchange)
    \item Weights: Pre-loaded from \texttt{backend/checkpoints/model.pth}
    \item Model path: \texttt{backend/checkpoints/efficientnet.onnx}
    \item Conversion: Uses onnx2pytorch library to convert ONNX to PyTorch format
    \item Input: 256×256×3 (height, width, channels)
    \item Output: 2 logits (real class logit, fake class logit)
    \item Inference device: CPU or GPU (auto-selected)
\end{enumerate}

\section{Score Aggregation}

After processing all sampled frames:

\begin{enumerate}
    \item Extract real class logits from all frames → \texttt{real\_scores}
    \item Extract fake class logits from all frames → \texttt{fake\_scores}
    \item Compute mean of each:
    \begin{equation}
    \text{real\_mean} = \frac{1}{n}\sum_{i=1}^{n} \text{real\_scores}[i]
    \end{equation}
    \begin{equation}
    \text{fake\_mean} = \frac{1}{n}\sum_{i=1}^{n} \text{fake\_scores}[i]
    \end{equation}
\end{enumerate}

\section{Final Decision Logic}

\begin{lstlisting}[language=Python]
if real_mean > 0.5:
    verdict = "REAL"
    confidence = real_mean * 100  # Convert to percentage
else:
    verdict = "FAKE"
    confidence = fake_mean * 100
\end{lstlisting}

The threshold of 0.5 is hardcoded in the logit space. No dynamic adjustment is performed for video.

\section{Output Format}

The video detection returns:

\begin{lstlisting}[language=Python]
{
    "label": "REAL" or "FAKE",
    "result": f"The video is {label}. Confidence: {confidence}%",
    "confidence": float  # 0-100 percentage
}
\end{lstlisting}

\section{Limitations}

\begin{itemize}
    \item Only 3 frames sampled (may miss localized DeepFakes)
    \item No temporal consistency check
    \item Simple averaging may underestimate frame-level artifacts
    \item No audio analysis (voice deepfakes)
    \item Fallback to image detection not fully implemented
\end{itemize}

% ===============================================
\chapter{Backend API Specification}
\label{ch:api}
% ===============================================

\section{Flask Application Structure}

File location: \texttt{backend/app.py}

The Flask application provides both server-side rendered pages and RESTful API endpoints.

\section{Health Check Endpoint}

\subsection{GET /api/health}

Returns system status and model availability.

\textbf{Response} (200 OK):

\begin{lstlisting}[language=Python]
{
    "success": True,
    "status": "ok",
    "python_executable": "/path/to/python",
    "python_version": "3.13.3",
    "runtime_initialized": True or False,
    "ai_init_mode": "async",
    "models": {
        "ai_predictor_loaded": bool,
        "ai_predictor_initialized": bool,
        "ai_predictor_init_in_progress": bool,
        "ai_predictor_error": null or "error message",
        "synthid_loaded": bool,
        "synthid_initialized": bool,
        "synthid_init_in_progress": bool,
        "synthid_error": null or "error message"
    },
    "c2pa": {
        "available": bool,
        "import_error": null or "error message",
        "has_reader": bool,
        "has_c2pa_error": bool
    }
}
\end{lstlisting}

\section{Image Analysis Endpoint}

\subsection{POST /api/analyze}

Analyzes an uploaded image through the complete three-layer detection pipeline.

\textbf{Request}:
\begin{itemize}
    \item Content-Type: multipart/form-data
    \item Field: \texttt{file} (image file)
    \item Allowed formats: PNG, JPG, JPEG, WEBP
    \item Max size: 16 MB
\end{itemize}

\textbf{Processing Steps}:

\begin{enumerate}
    \item Validate file format and size
    \item Save file to temporary upload directory
    \item Sleep 1.5 seconds for UX feedback
    \item Perform Layer 1: C2PA check
    \item Sleep 1.0 seconds
    \item Perform Layer 2: SynthID detection
    \item Sleep 0.5 seconds (if SynthID skipped)
    \item Sleep 2.0 seconds before AI model
    \item Perform Layer 3: AI ensemble + forensics
    \item Generate forensic report
    \item Clean up temporary file
    \item Return result
\end{enumerate}

The deliberate delays ensure frontend progress indicators have time to update.

\textbf{Response} (200 OK):

\begin{lstlisting}[language=Python]
{
    "success": True,
    "filename": "image_name.jpg",
    "layers": {
        "c2pa": {
            "status": "not_found|present|verified|unavailable",
            "c2pa_present": bool,
            "available": bool,
            "valid": bool,
            "issuer": null or "issuer_name",
            "ai_generated": bool,
            "raw_data": null or {...}
        },
        "synthid": {
            "status": "complete|skipped|unavailable|error",
            "available": bool,
            "is_watermarked": bool,
            "is_watermarked_raw": bool,
            "message": "string",
            "confidence": float,  # 0-100
            "raw_confidence": float,  # 0-1
            "correlation": float,
            "phase_match": float,
            "structure_ratio": float,
            "carrier_strength": float,
            "carrier_match_ratio": float,
            "multi_scale_consistency": float,
            "correlation_margin": float,
            "signal_score": float,
            "signal_band": "strong|weak|negative",
            "threshold": float or null,
            "calibration": {
                "applied": bool,
                "min_confidence": float,
                "min_phase_match": float,
                "min_correlation_margin": float,
                "min_carrier_match_ratio": float,
                "min_signal_score": float,
                "structure_ratio_range": [float, float],
                "max_multi_scale_std": float
            },
            "source": "reverse-SynthID robust extractor"
        },
        "ai_model": {
            "status": "complete|skipped|error",
            "label": "AI Image|Real Image",
            "confidence": float,  # 0-100
            "ai_probability": float,  # 0-100
            "source": "string",
            "model_scores": {
                "resnet_ai_score": float,
                "vit_ai_score": float or null,
                "legacy_ensemble_ai_score": float,
                "custom_forensic_ai_score": float,
                "forensics_only_ai_score": float,
                "weighted_custom_ai_score": float,
                "final_ai_score": float,
                "ai_decision_threshold": float,
                "base_blend_ai_score": float,
                "meta_ai_score": float or null,
                "vit_guard_applied": bool,
                "real_guard_applied": bool,
                "score_consensus_ai_score": float,
                "score_consensus_real_score": float,
                "score_margin": float,
                "score_consensus_refinement": bool
            },
            "forensic_modules": {
                "frequency": {
                    "available": bool,
                    "ai_score": float,
                    "detected": bool,
                    "message": "string",
                    "symmetric_peak_pairs": int,
                    "grid_energy_ratio": float,
                    "peak_prominence": float
                },
                "anatomy": {
                    "available": bool,
                    "ai_score": float,
                    "detected": bool,
                    "message": "string",
                    "details": {...}
                },
                "entropy": {
                    "available": bool,
                    "ai_score": float,
                    "detected": bool,
                    "message": "string",
                    "global_entropy": float,
                    "center_entropy": float,
                    "edge_entropy": float,
                    "local_entropy_spread": float,
                    "low_entropy_fraction": float
                }
            },
            "pipeline_override": {
                "applied": bool,
                "reason": null or "string"
            }
        }
    },
    "explainability": {
        "artifacts": ["artifact_description_1", "artifact_description_2"]
    },
    "forensic_summary": {
        "method": "string",
        "points": ["point_1", "point_2"],
        "conclusion": "string"
    },
    "final_verdict": "AI Generated (C2PA Metadata Present)|
                      AI Generated (SynthID Watermark Detected)|
                      AI Image|Real Image|Unknown",
    "confidence": float,  # 0-100
    "is_ai_generated": bool
}
\end{lstlisting}

\textbf{Error Response} (400/500):

\begin{lstlisting}[language=Python]
{
    "success": False,
    "error": "error message string"
}
\end{lstlisting}

\section{Video Analysis Endpoint}

\subsection{POST /api/analyze\_video}

Analyzes an uploaded video file.

\textbf{Request}:
\begin{itemize}
    \item Content-Type: multipart/form-data
    \item Field: \texttt{file} (video file)
    \item Allowed formats: MP4, AVI, MOV
    \item Max size: 16 MB
\end{itemize}

\textbf{Response} (200 OK):

\begin{lstlisting}[language=Python]
{
    "success": True,
    "label": "REAL|FAKE",
    "result": "The video is REAL/FAKE. Confidence score: X.X%",
    "confidence": float,  # 0-100
    "forensic_summary": {
        "method": "string",
        "points": [...],
        "conclusion": "string"
    }
}
\end{lstlisting}

\section{Forensic Report Endpoint}

\subsection{POST /api/forensic-report}

Generates detailed forensic report from analysis results.

\textbf{Request}:

\begin{lstlisting}[language=Python]
{
    "analysis_result": {...}  # Full response from /api/analyze or video
}
\end{lstlisting}

\textbf{Response} (200 OK): Detailed forensic report with enhanced insights.

\section{Page Routes}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Flask Page Routes}
\begin{tabular}{|l|l|p{4cm}|}
\hline
\textbf{Route} & \textbf{Method} & \textbf{Purpose} \\
\hline
\texttt{/} & GET & Home/landing page \\
\hline
\texttt{/dashboard} & GET & Image analysis UI \\
\hline
\texttt{/video\_dashboard} & GET & Video analysis UI \\
\hline
\texttt{/report} & GET & Image forensic report page \\
\hline
\texttt{/video\_report} & GET & Video forensic report page \\
\hline
\end{tabular}
\end{table}

% ===============================================
\chapter{Frontend Architecture}
\label{ch:frontend}
% ===============================================

\section{Technology Stack}

\begin{itemize}
    \item React 18.2.0 - UI framework
    \item Vite 5.1.0 - Build tool and dev server
    \item React Router 6.22.0 - Client-side routing
    \item TailwindCSS 3.4.1 - Utility-first CSS
    \item PostCSS - CSS preprocessing
\end{itemize}

\section{Project Structure}

\begin{lstlisting}[language=bash]
frontend/
├── src/
│   ├── App.jsx                    # Main router component
│   ├── main.jsx                   # Entry point
│   ├── index.css                  # Global styles
│   └── pages/
│       ├── Home.jsx               # Landing page
│       ├── Dashboard.jsx          # Image analysis UI
│       ├── Report.jsx             # Image report page
│       ├── VideoDashboard.jsx     # Video analysis UI
│       └── VideoReport.jsx        # Video report page
├── vite.config.js                 # Vite configuration
├── tailwind.config.js             # Tailwind CSS config
├── postcss.config.js              # PostCSS config
└── package.json                   # Dependencies
\end{lstlisting}

\section{Pages and Components}

\subsection{Home Page (Home.jsx)}

Landing page with entry points for image and video analysis.

\begin{itemize}
    \item Displays two main cards: ``Scan Image'' and ``Scan Video''
    \item System status indicator
    \item Navigation links to dashboard pages
    \item Dark cyberpunk aesthetic with neon accents
\end{itemize}

\subsection{Dashboard (Dashboard.jsx)}

Image file upload and real-time analysis progress.

\textbf{Features}:
\begin{itemize}
    \item Drag-and-drop file upload area
    \item Real-time progress indicator:
    \begin{enumerate}
        \item C2PA Metadata check
        \item SynthID Watermark detection
        \item Deep Learning Analysis
    \end{enumerate}
    \item Live console log showing step-by-step progress
    \item File information display
    \item Error handling and user feedback
\end{itemize}

\textbf{API Calls}:
\begin{enumerate}
    \item GET \texttt{/api/health} - Check system ready
    \item POST \texttt{/api/analyze} - Perform image analysis
    \item POST \texttt{/api/forensic-report} - Generate detailed report
\end{enumerate}

\subsection{Report (Report.jsx)}

Detailed forensic analysis results for analyzed image.

\textbf{Displays}:
\begin{itemize}
    \item Large confidence gauge showing prediction
    \item Final verdict (AI Image / Real Image)
    \item Detailed breakdown by analysis layer:
    \begin{itemize}
        \item C2PA metadata status
        \item SynthID watermark detection results
        \item AI model scores and confidence
        \item Forensic module results (frequency, anatomy, entropy)
    \end{itemize}
    \item Explainability: List of detected artifacts
    \item Full forensic summary with conclusions
\end{itemize}

\subsection{VideoDashboard (VideoDashboard.jsx)}

Video file upload and analysis.

\textbf{Similar to Dashboard but}:
\begin{itemize}
    \item Accepts video formats (MP4, AVI, MOV)
    \item Larger file size tolerance
    \item Progress shows: Frame Sampling → Video Analysis
\end{itemize}

\subsection{VideoReport (VideoReport.jsx)}

Video analysis results.

\textbf{Displays}:
\begin{itemize}
    \item Video verdict (REAL / FAKE)
    \item Confidence percentage
    \item Forensic summary
\end{itemize}

\section{Styling and Theme}

\textbf{Overall Design}: Dark cyberpunk aesthetic

\begin{itemize}
    \item Primary color: Neon cyan (#00f2ff)
    \item Secondary color: Neon green (#00ff88)
    \item Background: Dark grayscale with dark blue tints
    \item Glass-morphism cards: Frosted glass effect with backdrop blur
    \item Gradient text: Cyan-to-blue gradients for emphasis
    \item Animations: Fade-in, slide-in, glow effects
\end{itemize}

\textbf{Key CSS Classes}:

\begin{lstlisting}[language=CSS]
.glass-card {
    background: rgba(255, 255, 255, 0.1);
    backdrop-filter: blur(10px);
    border: 1px solid rgba(255, 255, 255, 0.2);
}

.gradient-text {
    background: linear-gradient(to right, #00f2ff, #0088ff);
    -webkit-background-clip: text;
    -webkit-text-fill-color: transparent;
}

.btn-neon {
    background: #00f2ff;
    color: #000;
    box-shadow: 0 0 10px #00f2ff;
}

.btn-neon:hover {
    box-shadow: 0 0 20px #00f2ff;
}
\end{lstlisting}

\section{Development Server Configuration}

\textbf{Vite Config (vite.config.js)}:

\begin{lstlisting}[language=JavaScript]
import { defineConfig } from 'vite'
import react from '@vitejs/plugin-react'

export default defineConfig({
    plugins: [react()],
    server: {
        proxy: {
            '/api': {
                target: 'http://localhost:7860',
                changeOrigin: true
            }
        }
    }
})
\end{lstlisting}

The proxy configuration allows the frontend dev server (running on port 5173) to proxy API requests to the Flask backend (port 7860) during development.

% ===============================================
\chapter{Model Weights and Configuration}
\label{ch:models}
% ===============================================

\section{Model Files and Locations}

All model weights are downloaded from HuggingFace repository on first run.

\subsection{Image Analysis Models}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Image Analysis Model Files}
\begin{tabular}{|l|l|r|}
\hline
\textbf{File} & \textbf{Location} & \textbf{Size} \\
\hline
ResNet50 weights & \texttt{model\_output/resnet50\_finetuned\_benchmark.pth} & $\sim$100 MB \\
\hline
Meta-learner & \texttt{model\_output/ai\_detector\_meta\_learner.joblib} & $\sim$1-5 MB \\
\hline
Polynomial transformer & \texttt{model\_output/polynomial\_transformer.joblib} & $<$1 MB \\
\hline
SynthID codebook & \texttt{model\_output/synthid/robust\_codebook.pkl} & $\sim$50-100 MB \\
\hline
ViT (HuggingFace) & \texttt{dima806/ai\_vs\_real\_image\_detection} & $\sim$350 MB \\
\hline
\end{tabular}
\end{table}

\subsection{Video Analysis Models}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Video Analysis Model Files}
\begin{tabular}{|l|l|r|}
\hline
\textbf{File} & \textbf{Location} & \textbf{Size} \\
\hline
EfficientNet ONNX & \texttt{backend/checkpoints/efficientnet.onnx} & $\sim$100 MB \\
\hline
EfficientNet weights & \texttt{backend/checkpoints/model.pth} & $\sim$200 MB \\
\hline
\end{tabular}
\end{table}

\section{Model Download Mechanism}

Models are downloaded from HuggingFace repository \texttt{Adhil786/deepfake-models} on server startup if not present locally.

\textbf{Implementation in app.py}:

\begin{lstlisting}[language=Python]
from huggingface_hub import hf_hub_download
import os
from pathlib import Path

repo = "Adhil786/deepfake-models"
token = os.environ.get("HF_TOKEN", None)

files_to_download = [
    "model_output/synthid/robust_codebook.pkl",
    "model_output/resnet50_finetuned_benchmark.pth",
    "ai_detector_meta_learner.joblib",
    "polynomial_transformer.joblib",
]

for filepath in files_to_download:
    full_path = os.path.join(project_root, filepath)
    if not Path(full_path).exists():
        hf_hub_download(
            repo_id=repo,
            filename=filepath,
            local_dir=project_root,
            token=token
        )
\end{lstlisting}

\section{Environment Variables for Model Configuration}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Model-Related Environment Variables}
\begin{tabular}{|l|l|p{5cm}|}
\hline
\textbf{Variable} & \textbf{Default} & \textbf{Meaning} \\
\hline
\texttt{RESNET\_AI\_INDEX} & 1 & Output neuron index for AI class in ResNet \\
\hline
\texttt{VIT\_AI\_INDEX} & 1 (auto-detect) & Output neuron index for AI class in ViT \\
\hline
\texttt{ENABLE\_TTA} & 1 & Enable test-time augmentation (horizontal flip) \\
\hline
\texttt{SKIP\_VIT} & 0 & Skip ViT model loading (for slow environments) \\
\hline
\texttt{HF\_TOKEN} & (none) & HuggingFace API token for authenticated downloads \\
\hline
\end{tabular}
\end{table}

% ===============================================
\chapter{All Thresholds and Parameters}
\label{ch:thresholds}
% ===============================================

\section{SynthID Watermark Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{SynthID Detection Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Meaning} \\
\hline
\texttt{SYNTHID\_MIN\_CONFIDENCE} & 0.50 & Minimum raw watermark confidence (0-1) \\
\hline
\texttt{SYNTHID\_MIN\_PHASE\_MATCH} & 0.40 & Minimum phase alignment (0-1) \\
\hline
\texttt{SYNTHID\_MIN\_CORRELATION\_MARGIN} & 0.0001 & Minimum correlation margin \\
\hline
\texttt{SYNTHID\_MIN\_CARRIER\_MATCH\_RATIO} & 0.15 & Minimum carrier match (0-1) \\
\hline
\texttt{SYNTHID\_MIN\_SIGNAL\_SCORE} & 0.45 & Minimum signal quality (0-1) \\
\hline
\texttt{SYNTHID\_MIN\_STRUCTURE\_RATIO} & 0.78 & Minimum structure ratio \\
\hline
\texttt{SYNTHID\_MAX\_STRUCTURE\_RATIO} & 1.85 & Maximum structure ratio \\
\hline
\texttt{SYNTHID\_MAX\_MULTI\_SCALE\_STD} & 0.30 & Maximum inconsistency (0-1) \\
\hline
\end{tabular}
\end{table}

\section{Frequency Analysis Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Frequency Module Detection Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Meaning} \\
\hline
Detection threshold & 0.55 & AI score must be $\geq$ 0.55 \\
\hline
Peak count threshold & 4 & Minimum symmetric peaks required \\
\hline
Grid energy ratio & 1.18 & Minimum grid energy ratio \\
\hline
Search radius (norm) & 0.22-0.88 & Normalized freq space region \\
\hline
\end{tabular}
\end{table}

\section{Anatomy Analysis Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Anatomy Analysis Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Meaning} \\
\hline
Limb ratio anomaly & 1.55x & Ratio indicating anomaly \\
\hline
Limb ratio hard limit & 2.10x & Extreme limit \\
\hline
Joint angle min & 8° & Minimum natural joint angle \\
\hline
Joint angle max & 178° & Maximum natural joint angle \\
\hline
Finger segment ratio & 1.95x & Normal ratio bound \\
\hline
Finger ratio hard limit & 3.20x & Extreme limit \\
\hline
Hand width ratio & 0.45-2.40x & Hand shape bounds \\
\hline
\end{tabular}
\end{table}

\section{Entropy Analysis Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Entropy Analysis Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Meaning} \\
\hline
Detection threshold & 0.45 & AI score must be $\geq$ 0.45 \\
\hline
Center gap threshold & 0.60 & Minimum center-edge entropy gap \\
\hline
Entropy spread threshold & 1.10 & Minimum local variation \\
\hline
Low entropy fraction & 0.35 & Fraction of bland patches \\
\hline
Guard: low spread & 0.45 & Reset bland penalty if < this \\
\hline
Guard: low gap & 0.15 & Reset bland penalty if center gap < this \\
\hline
\end{tabular}
\end{table}

\section{AI Probability Decision Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Dynamic Decision Thresholds}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Forensic Support Level} & \textbf{ViT Available} & \textbf{ResNet Only} \\
\hline
2+ modules $\geq 0.60$ & 0.54 & 0.57 \\
\hline
1 module strong $\geq 0.60$ & 0.58 & 0.62 \\
\hline
1 module weak 0.35-0.60 & 0.62 & 0.66 \\
\hline
No support, 0 negatives & 0.64 & 0.68 \\
\hline
No support, 1 negative $\leq 0.15$ & 0.70-0.72 & 0.74-0.76 \\
\hline
No support, 2+ negatives & 0.76-0.80 & 0.80+ \\
\hline
\end{tabular}
\end{table}

\section{Forensic Weighting}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Forensic Score Weights}
\begin{tabular}{|l|r|}
\hline
\textbf{Component} & \textbf{Weight} \\
\hline
ResNet50 & 50\% \\
\hline
Frequency analysis & 20\% \\
\hline
Anatomy analysis & 20\% \\
\hline
Entropy analysis & 10\% \\
\hline
\textbf{Total} & \textbf{100\%} \\
\hline
\end{tabular}
\end{table}

\section{Support/Negative Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Evidence Classification Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Classification} & \textbf{Score} & \textbf{Meaning} \\
\hline
Support evidence & $\geq 0.60$ & Strong signal for AI \\
\hline
Neutral & 0.15 – 0.60 & Ambiguous \\
\hline
Negative evidence & $\leq 0.15$ & Signal against AI \\
\hline
Forensic support threshold & 0.60 & Support level \\
\hline
Forensic negative threshold & 0.15 & Negative level \\
\hline
\end{tabular}
\end{table}

\section{Video Detection Thresholds}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Video Detection Thresholds}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Meaning} \\
\hline
Frames to sample & 3 & Frame count from video \\
\hline
Decision threshold & 0.5 & Real/fake cutoff in logit space \\
\hline
Frame positions & 0\%, 50\%, 100\% & Sampling distribution \\
\hline
Frame resize & 256×256 & Input dimension \\
\hline
Normalization divisor & 255.0 & Pixel value scaling \\
\hline
\end{tabular}
\end{table}

% ===============================================
\chapter{Configuration and Environment Variables}
\label{ch:configuration}
% ===============================================

\section{Complete Environment Variable Reference}

\subsection{Server Configuration}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Flask Server Configuration}
\begin{tabular}{|l|l|p{4cm}|}
\hline
\textbf{Variable} & \textbf{Default} & \textbf{Purpose} \\
\hline
\texttt{PORT} & 7860 & Server port number \\
\hline
\texttt{HOST} & 0.0.0.0 & Server host address \\
\hline
\texttt{WAITRESS\_THREADS} & 8 & Production server thread count \\
\hline
\texttt{DEBUG} & False & Flask debug mode \\
\hline
\end{tabular}
\end{table}

\subsection{Model Configuration}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Model Loading Configuration}
\begin{tabular}{|l|l|p{4cm}|}
\hline
\textbf{Variable} & \textbf{Default} & \textbf{Purpose} \\
\hline
\texttt{RESNET\_AI\_INDEX} & 1 & ResNet AI output index \\
\hline
\texttt{VIT\_AI\_INDEX} & auto & ViT AI output index \\
\hline
\texttt{ENABLE\_TTA} & 1 & Test-time augmentation \\
\hline
\texttt{SKIP\_VIT} & 0 & Skip ViT loading \\
\hline
\texttt{HF\_TOKEN} & (none) & HuggingFace API token \\
\hline
\end{tabular}
\end{table}

\subsection{SynthID Configuration}

Covered in Chapter \ref{ch:thresholds}.

\subsection{MediaPipe Configuration}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{MediaPipe Model Paths}
\begin{tabular}{|l|l|p{4cm}|}
\hline
\textbf{Variable} & \textbf{Default} & \textbf{Purpose} \\
\hline
\texttt{MEDIAPIPE\_POSE\_MODEL\_PATH} & auto & Pose landmarker model file \\
\hline
\texttt{MEDIAPIPE\_HAND\_MODEL\_PATH} & auto & Hand landmarker model file \\
\hline
\end{tabular}
\end{table}

% ===============================================
\chapter{Data Flow Diagrams and Algorithms}
\label{ch:dataflow}
% ===============================================

\section{Complete Image Analysis Flow}

\begin{enumerate}
    \item User uploads image through web interface
    \item Frontend sends POST request to \texttt{/api/analyze}
    \item Backend validates file format and size
    \item File saved to temporary upload directory
    \item \textbf{Layer 1: C2PA Check}
    \begin{itemize}
        \item Read C2PA manifest from image
        \item If present: Return "AI Generated" immediately
        \item If absent: Continue
    \end{itemize}
    \item \textbf{Layer 2: SynthID Detection}
    \begin{itemize}
        \item Extract watermark using robust extractor
        \item Apply calibration logic
        \item If watermark detected: Return "AI Generated"
        \item If not detected: Continue
    \end{itemize}
    \item \textbf{Layer 3: Deep Learning Ensemble}
    \begin{itemize}
        \item Inference with ResNet50
        \item Inference with ViT (optional)
        \item Combine scores through meta-learner
        \item Apply forensic analysis (frequency, anatomy, entropy)
        \item Apply guards and checks
        \item Determine dynamic threshold
        \item Make final AI/Real decision
    \end{itemize}
    \item Generate forensic report
    \item Compile response JSON
    \item Clean up temporary file
    \item Return response to frontend
    \item Frontend displays verdict and forensic details
\end{enumerate}

\section{SynthID Watermark Extraction Algorithm}

\begin{enumerate}
    \item Input: Image file
    
    \item Multi-scale processing (3 scales: 256, 512, 1024)
    \begin{itemize}
        \item Resize image to each scale
        \item Process each scale independently
        \item Fuse results at end
    \end{itemize}
    
    \item For each scale:
    \begin{itemize}
        \item Apply 4 denoising methods (wavelet, bilateral, NLM, Wiener)
        \item Fuse denoised results
        \item Compute FFT and power spectrum
        \item Search for known SynthID carrier frequencies
        \item Measure phase alignment
        \item Estimate correlation margin above noise
    \end{itemize}
    
    \item Apply ICA/PCA decomposition to separate watermark from image
    
    \item Compute composite signal score from:
    \begin{itemize}
        \item Raw confidence
        \item Phase match
        \item Structure ratio validity
        \item Multi-scale consistency
        \item Carrier match strength
    \end{itemize}
    
    \item Apply calibration logic:
    \begin{itemize}
        \item If raw detector fired: Relax thresholds
        \item If raw didn't fire: Apply strict thresholds
    \end{itemize}
    
    \item Return: is\_watermarked boolean + detailed scores
\end{enumerate}

\section{AI Score Fusion Algorithm}

\begin{enumerate}
    \item Get ResNet50 output → \texttt{resnet\_score}
    
    \item Get ViT output (optional) → \texttt{vit\_score}
    
    \item If ViT available:
    \begin{itemize}
        \item Compute base blend: $(resnet + vit) / 2$
        \item Load meta-learner and polynomial transformer
        \item Create polynomial features from [resnet, vit]
        \item Meta-learner prediction → \texttt{meta\_score}
        \item Check drift between blend and meta
        \item If drift > threshold: Average them
        \item Otherwise: Use meta\_score
    \end{itemize}
    
    \item If ViT unavailable:
    \begin{itemize}
        \item Use ResNet score as baseline
    \end{itemize}
    
    \item Run forensic modules (frequency, anatomy, entropy)
    
    \item Classify forensic support level (0, 1, 2+)
    
    \item Apply forensic calibration rules to adjust final score
    
    \item Apply guards:
    \begin{itemize}
        \item ViT disagreement guard (if activates)
        \item Real disagreement guard (if activates)
    \end{itemize}
    
    \item Determine dynamic decision threshold based on:
    \begin{itemize}
        \item Forensic support count
        \item Negative evidence count
        \item Model disagreement level
        \item Feature confidence margins
    \end{itemize}
    
    \item Compare final AI score to threshold
    
    \item Return: Verdict (AI/Real) + Confidence (0-100)
\end{enumerate}

\section{Frequency Artifact Detection Algorithm}

\begin{enumerate}
    \item Input: Image
    
    \item Convert to grayscale
    
    \item Apply Hanning window (for FFT preparation)
    
    \item Compute 2D FFT
    
    \item Shift zero-freq to center
    
    \item Calculate power spectral density in log scale
    
    \item Focus on annular region (radius 0.22-0.88 norm)
    
    \item Search for:
    \begin{itemize}
        \item Symmetric peak pairs
        \item Grid alignment energy
        \item Harmonic patterns
    \end{itemize}
    
    \item Measure baseline noise level
    
    \item Score features:
    \begin{itemize}
        \item Peak prominence
        \item Symmetric pair count
        \item Grid energy ratio
        \item Axis alignment
    \end{itemize}
    
    \item Compute weighted AI score
    
    \item Check against 0.55 threshold OR (4+ peaks AND ratio > 1.18)
    
    \item Output: AI score + detected boolean
\end{enumerate}

% ===============================================
\chapter{Error Handling and Fallback Behaviors}
\label{ch:errors}
% ===============================================

\section{Model Unavailability}

\subsection{ViT Model Missing}
If ViT cannot be downloaded or loaded:
\begin{itemize}
    \item System continues to work with ResNet50 only
    \item Score fusion reverts to averaging or single-model output
    \item Confidence may be slightly lower
    \item No error shown to user
\end{itemize}

\subsection{SynthID Module Missing}
If SynthID is unavailable:
\begin{itemize}
    \item Layer 2 is skipped
    \item Process continues to Layer 3 (deep learning)
    \item No error message
    \item Final verdict still made
\end{itemize}

\subsection{Anatomy Forensic Module Missing}
If MediaPipe or anatomy analysis fails:
\begin{itemize}
    \item Forensic module returns ``unavailable''
    \item Remaining forensic modules proceed
    \item Decision threshold adjusted upward (more conservative)
    \item Special guard activated: \texttt{threshold = 0.86}
\end{itemize}

\section{File Upload Errors}

\begin{itemize}
    \item Invalid format: 400 error
    \item File too large: 400 error
    \item Corrupted file: 500 error with error message
    \item Temporary file cleanup always executed
\end{itemize}

\section{Network Errors}

\begin{itemize}
    \item Model download failure: Displayed at startup
    \item C2PA library import failure: Gracefully noted
    \item Video processing timeout: Returns error response
\end{itemize}

% ===============================================
\chapter{Complete Technical Summary}
\label{ch:summary}
% ===============================================

\section{Key Technical Achievements}

\begin{enumerate}
    \item \textbf{Multi-Layer Detection}: Three independent detection layers provide defense in depth
    
    \item \textbf{SynthID Watermark Detection}: Implements sophisticated multi-scale, multi-denoiser watermark extraction
    
    \item \textbf{AI Ensemble}: Combines ResNet50 and Vision Transformer through trained meta-learner
    
    \item \textbf{Forensic Analysis}: Three forensic modules (frequency, anatomy, entropy) provide independent evidence
    
    \item \textbf{Dynamic Thresholds}: Decision threshold adapts based on forensic support and model agreement
    
    \item \textbf{Intelligent Guards}: Corrects model disagreement using score consensus algorithms
    
    \item \textbf{Gradual Score Combination}: Blends multiple scoring methods with checks for drift and disagreement
    
    \item \textbf{Comprehensive Calibration}: Every parameter is environment variable configurable
\end{enumerate}

\section{Unique Features}

\begin{itemize}
    \item \textbf{C2PA Support}: Checks for industry-standard content authenticity metadata
    
    \item \textbf{Multi-Denoiser Fusion}: Uses 4 denoising methods to improve watermark extraction
    
    \item \textbf{Known Carrier Detection}: Pre-computed database of SynthID carrier frequencies from 288 samples
    
    \item \textbf{Forensic Cross-Check}: Independent forensic evidence prevented false positives
    
    \item \textbf{Test-Time Augmentation}: Optional horizontal flip during inference for robustness
    
    \item \textbf{Explainability}: Reports list of detected artifacts
\end{itemize}

\section{Performance Characteristics}

\begin{itemize}
    \item Image analysis: ~5-10 seconds end-to-end (with deliberate delays for UX)
    \item Video analysis: ~10-15 seconds (3 frames + processing)
    \item Model memory: ~1-2 GB peak (with all models loaded)
    \item Inference device: Auto-selects GPU if available, falls back to CPU
\end{itemize}

\section{Deployment Considerations}

\begin{itemize}
    \item Frontend: Requires Node.js 16+ and npm for building
    \item Backend: Requires Python 3.10+
    \item Production: Waitress WSGI server with 8 threads
    \item Models: ~750 MB total downloaded on first run
    \item Port: Flask server runs on port 7860 by default
    \item Proxy: Frontend dev server proxies /api to backend
\end{itemize}

% ===============================================
\chapter{Deployment and Design Decisions}
\label{ch:deployment}
% ===============================================

\section{Production Deployment}

\subsection{Server Configuration}

The system uses Waitress WSGI server for production:

\begin{lstlisting}[language=Python]
from waitress import serve
from backend.app import app

serve(
    app,
    host="0.0.0.0",
    port=7860,
    threads=8  # Thread pool size
)
\end{lstlisting}

This multi-threaded configuration allows 8 concurrent requests per instance.

\subsection{Environment Requirements}

\textbf{Required}:
\begin{itemize}
    \item Python 3.10, 3.11, 3.12, or 3.13
    \item Node.js 18 or higher (for frontend building)
    \item 2+ GB RAM minimum
    \item 1+ GB disk space for models
\end{itemize}

\textbf{Optional}:
\begin{itemize}
    \item NVIDIA GPU with CUDA for accelerated inference
    \item HuggingFace API token for authenticated model downloads
\end{itemize}

\subsection{Windows Startup Scripts}

Located in \texttt{scripts/windows/}:

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Windows Startup Scripts}
\begin{tabular}{|l|p{5cm}|}
\hline
\textbf{Script} & \textbf{Purpose} \\
\hline
\texttt{start-backend-dev.ps1} & Development server with Flask auto-reload \\
\hline
\texttt{start-backend-prod.ps1} & Production server with Waitress \\
\hline
\texttt{start-frontend-dev.ps1} & Frontend development server on port 5173 \\
\hline
\texttt{health-check.ps1} & Verify backend is responsive and ready \\
\hline
\end{tabular}
\end{table}

\subsection{Initialization Sequence}

When the Flask application starts:

\begin{enumerate}
    \item Flask app object created and configured
    \item HTTP server binds to port 7860
    \item Background initialization thread spawned:
    \begin{enumerate}
        \item Check for required model files in \texttt{model\_output/}
        \item Download missing files from HuggingFace repo \texttt{Adhil786/deepfake-models}
        \item Load ResNet50 model (required, must succeed)
        \item Load Vision Transformer (optional, skipped if \texttt{SKIP\_VIT=1})
        \item Load meta-learner and polynomial transformer
        \item Load SynthID service
        \item Update initialization status in memory
    \end{enumerate}
    \item API endpoints become responsive immediately (before background initialization completes)
    \item \texttt{GET /api/health} endpoint reports initialization progress
    \item Client receives response "models in progress" if still loading
\end{enumerate}

This asynchronous loading allows the server to start quickly while models load in the background.

\subsection{Model Download Mechanism}

When required model files are missing locally, they are automatically downloaded from HuggingFace:

\begin{lstlisting}[language=Python]
from huggingface_hub import hf_hub_download
import os

repo = "Adhil786/deepfake-models"
token = os.environ.get("HF_TOKEN", None)

files = [
    "model_output/synthid/robust_codebook.pkl",
    "model_output/resnet50_finetuned_benchmark.pth",
    "ai_detector_meta_learner.joblib",
    "polynomial_transformer.joblib"
]

for filepath in files:
    full_path = os.path.join(project_root, filepath)
    if not os.path.exists(full_path):
        hf_hub_download(
            repo_id=repo,
            filename=filepath,
            local_dir=project_root,
            token=token
        )
\end{lstlisting}

Total download size: Approximately 750 MB across all models.

\section{Notable Design Decisions}

\subsection{Decision 1: Cascade Architecture with Short-Circuit Logic}

\textbf{Design}: Three-layer cascade where successful detection at Layer 1 or 2 stops further analysis.

\textbf{Rationale}:
\begin{itemize}
    \item C2PA metadata is strongest possible evidence (industry standard authentication)
    \item SynthID watermark is high-confidence signal from Google
    \item Proceeding to deep learning when high-confidence signals exist wastes computation
    \item Early returns provide fast feedback for clear-cut cases
\end{itemize}

\subsection{Decision 2: SynthID Dual Detection Paths}

\textbf{Design}: Two different threshold sets depending on whether raw detector fired.

\textbf{Rationale}:
\begin{itemize}
    \item Raw detector tuned for maximum sensitivity in training environment
    \item Real-world images may differ from training data (environmental drift)
    \item If raw detector fires, indicates strong signal; relax secondary thresholds
    \item Prevents suppression of true positives due to environment differences
    \item False negative risk higher than false positive if signal initially detected
\end{itemize}

\subsection{Decision 3: Dynamic Threshold Based on Forensic Support}

\textbf{Design}: Decision threshold varies from 0.54 to 0.86 based on amount of forensic evidence.

\textbf{Rationale}:
\begin{itemize}
    \item When multiple forensic modules support AI verdict: Can use lower threshold (0.54)
    \item When forensic modules contradict AI verdict: Must use higher threshold (0.80+)
    \item Forensic evidence provides independent corroboration
    \item Balances confidence against risk of false positives
\end{itemize}

\subsection{Decision 4: Guards Against Model Disagreement}

\textbf{Design}: Special logic when ResNet and ViT strongly disagree.

\textbf{Rationale}:
\begin{itemize}
    \item ResNet and ViT learn different features (CNN vs Transformer)
    \item Strong disagreement indicates ambiguous case
    \item ViT may be learning spurious features unrelated to AI generation
    \item Guards check forensic support before trusting ViT
    \item Prevents ViT false positives when ResNet disagrees and forensics are weak
\end{itemize}

\subsection{Decision 5: Test-Time Augmentation (TTA)}

\textbf{Design}: Optional horizontal flip during inference for ResNet50.

\textbf{Rationale}:
\begin{itemize}
    \item Left-right symmetry in natural images
    \item AI generation may create asymmetric artifacts
    \item Averaging predictions from original and flipped improves robustness
    \item Cost: 2x CPU time but more stable predictions
    \item Disabled for speed if \texttt{ENABLE\_TTA=0}
\end{itemize}

\subsection{Decision 6: Graceful Degradation on Missing Models}

\textbf{Design}: System continues functioning if optional models unavailable.

\textbf{Rationale}:
\begin{itemize}
    \item ViT (~350 MB) may fail to download in bandwidth-limited environments
    \item ResNet (100 MB) sufficient for baseline accuracy
    \item Meta-learner may have version mismatches from retraining
    \item Fallback to averaging better than complete failure
    \item Critical models (ResNet, codebook) required; optional ones downgrade gracefully
\end{itemize}

\subsection{Decision 7: Deliberate Processing Delays}

\textbf{Design}: Intentional sleep() calls during processing (1.5s, 1.0s, 2.0s).

\textbf{Rationale}:
\begin{itemize}
    \item Provides user feedback opportunity in progress indicator
    \item Prevents server resource spikes from rapid requests
    \item Allows frontend to update progress UI smoothly
    \item Creates perceived sophistication of analysis
\end{itemize}

Delays can be removed for unattended batch processing if needed.

\subsection{Decision 8: Weighted Forensic Combination}

\textbf{Design}: Forensic modules weighted at 40\% total, ResNet dominates at 50\%.

\textbf{Rationale}:
\begin{itemize}
    \item Forensics are artifact detectors, not end-to-end classifiers
    \item Forensic false positives possible (e.g., JPEG compression looks like checkerboard)
    \item ResNet trained end-to-end on AI vs real classification task
    \item 50\% ResNet + 40\% forensics + 10\% consistency balances approaches
\end{itemize}

\section{Size Constraints and Performance}

\subsection{Upload Limits}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Upload Size Constraints}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Media Type} & \textbf{Max Size} & \textbf{Purpose} \\
\hline
Image upload & 16 MB & Temporary storage in \texttt{/uploads} \\
\hline
Video upload & 16 MB & Temporary storage in \texttt{/uploads} \\
\hline
Internal image resize & 512 px max side & Processing normalization \\
\hline
\end{tabular}
\end{table}

Limits configurable in Flask app via \texttt{MAX\_CONTENT\_LENGTH} setting.

\subsection{Inference Sequence Timing}

Complete request flow timing breakdown:

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Request Processing Timeline}
\begin{tabular}{|l|r|p{3cm}|}
\hline
\textbf{Stage} & \textbf{Duration} & \textbf{Purpose} \\
\hline
File upload and save & <1 second & Store tmp file \\
\hline
Delay & 1.5 seconds & UX feedback \\
\hline
C2PA check & 0.5 seconds & Read metadata \\
\hline
Delay & 1.0 seconds & UX feedback \\
\hline
SynthID watermark & 1-2 seconds & Multi-scale extraction \\
\hline
Delay & 0.5-2.0 seconds & UX feedback \\
\hline
ResNet50 inference & 0.5 seconds & Forward pass \\
\hline
ViT inference & 1-2 seconds & Transformer pass \\
\hline
Forensic analysis & 1-2 seconds & 3 modules \\
\hline
Score fusion & <0.1 seconds & Thresholds \\
\hline
Report generation & 0.5 seconds & Formatting \\
\hline
Cleanup & <0.1 seconds & Delete tmp file \\
\hline
\textbf{Total} & \textbf{5-10 seconds} & \textbf{Including delays} \\
\hline
\end{tabular}
\end{table}

\section{Resource Requirements}

\subsection{Memory Usage}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Estimated Memory Requirements}
\begin{tabular}{|l|r|p{4cm}|}
\hline
\textbf{Component} & \textbf{Memory} & \textbf{Notes} \\
\hline
ResNet50 model & 400 MB & Loaded once at startup \\
\hline
ViT model & 1.4 GB & Loaded once if available \\
\hline
Per-request overhead & 100-300 MB & Image + intermediate tensors \\
\hline
Meta-learner & 50 MB & Lightweight model \\
\hline
SynthID codebook & 200 MB & Hash tables and embeddings \\
\hline
Flask/Python runtime & 100 MB & Base interpreter \\
\hline
\textbf{Peak (all loaded)} & \textbf{2-2.5 GB} & \textbf{With ViT} \\
\hline
\textbf{Minimum (ResNet only)} & \textbf{700 MB} & \textbf{Without ViT} \\
\hline
\end{tabular}
\end{table}

\subsection{Disk Usage}

\begin{table}[H]
\centering
\captionsetup{justification=centering}
\caption{Disk Space Requirements}
\begin{tabular}{|l|r|}
\hline
\textbf{Component} & \textbf{Size} \\
\hline
ResNet50 weights & 100 MB \\
\hline
ViT model & 350 MB \\
\hline
SynthID codebook & 100 MB \\
\hline
Meta-learner + transformer & 10 MB \\
\hline
EfficientNet (video) & 300 MB \\
\hline
Frontend bundle & 500 KB \\
\hline
\textbf{Total} & \textbf{~860 MB} \\
\hline
\end{tabular}
\end{table}

\section{Example Complete Detection Flows}

\subsection{Example 1: AI Image with SynthID Watermark}

\textbf{Input}: Image generated by Google's image model with SynthID watermark

\begin{verbatim}
Layer 1 (C2PA):
  └─ No metadata found
     Continue to Layer 2

Layer 2 (SynthID):
  ├─ Multi-scale watermark extraction: SUCCESS
  ├─ Raw confidence: 0.72 (0-1 scale) = 72%
  ├─ Phase match: 0.55
  ├─ Signal score: 0.68
  ├─ Carrier match ratio: 0.35
  ├─ Multi-scale consistency: 0.25
  │
  ├─ Check calibrated decision:
  │  ├─ min_confidence (0.50): 0.72 ✓
  │  ├─ min_phase_match (0.40): 0.55 ✓
  │  ├─ min_signal_score (0.45): 0.68 ✓
  │  ├─ min_carrier_match (0.15): 0.35 ✓
  │  └─ All checks pass!
  │
  └─ DECISION: is_watermarked = TRUE

Final Verdict: "AI Generated (SynthID Watermark Detected)"
  │ Confidence: 95%
  │ Signal Band: STRONG
  └─ Layer 3 SKIPPED (short-circuit)
\end{verbatim}

\subsection{Example 2: Ambiguous Natural Photo with Weak Artifacts}

\textbf{Input}: Real photograph but with JPEG compression creating checkerboard patterns

\begin{verbatim}
Layer 1 (C2PA):
  └─ No metadata
     Continue

Layer 2 (SynthID):
  ├─ Watermark extraction: NOT FOUND
  ├─ Raw confidence: 0.38 (below 0.50 threshold)
  └─ is_watermarked = FALSE
     Continue to Layer 3

Layer 3 (Deep Learning):
  ├─ ResNet50: 0.35 confidence (Real)
  ├─ ViT: 0.38 confidence (Real)
  │
  ├─ Forensic Modules:
  │  ├─ Frequency: 0.62 (DETECTED - jpeg compression)
  │  ├─ Anatomy: 0.12 (negative)
  │  └─ Entropy: 0.08 (negative)
  │
  ├─ Support Level: 1 module (frequency only)
  │  └─ Apply Rule: "Be conservative with frequency"
  │  └─ final_ai_prob = 0.70*resnet + 0.30*weighted_avg
  │  └─ final_ai_prob = 0.70*0.35 + 0.30*0.40 = 0.365
  │
  ├─ Dynamic Threshold Calculation:
  │  ├─ Support: 1 (only frequency)
  │  ├─ Negatives: 2 (anatomy and entropy low)
  │  ├─ Disagreement: low (models agree on real)
  │  └─ threshold = 0.72 (conservative)
  │
  ├─ Final Decision:
  │  ├─ ai_prob (0.365) < threshold (0.72)
  │  └─ DECISION: Real Image
  │
  └─ Confidence: 60% real
     └─ Message: "Frequency analysis detected artifacts
                  typical of JPEG compression, but other
                  indicators support natural image."

Final Verdict: "Real Image"
  │ Confidence: 60%
  └─ Forensic notes with detailed explanation
\end{verbatim}

\subsection{Example 3: Weak Deepfake with ViT-ResNet Disagreement}

\textbf{Input}: Generated image with subtle artifacts, confuses Transformer

\begin{verbatim}
Layer 1 & 2: No signals (skip)

Layer 3 (Deep Learning):
  ├─ ResNet50: 0.32 (confident: REAL)
  ├─ ViT: 0.78 (confident: AI)
  │  └─ Strong disagreement! (0.78 - 0.32 = 0.46)
  │
  ├─ Forensic Modules:
  │  ├─ Frequency: 0.48 (borderline)
  │  ├─ Anatomy: 0.15 (neutral)
  │  └─ Entropy: 0.18 (weak)
  │  └─ ViT likely learning spurious features
  │
  ├─ Apply ViT Disagreement Guard:
  │  ├─ Activation check:
  │  │  ├─ ViT confident AI (0.78 ≥ 0.86)? NO
  │  │  └─ Guard does NOT activate
  │
  ├─ Base Blend:
  │  ├─ blend = (0.32 + 0.78) / 2 = 0.55
  │
  ├─ Meta-learner:
  │  ├─ Poly features from [0.32, 0.78]
  │  ├─ Prediction: 0.48 (AI prob)
  │  ├─ Drift: |0.48 - 0.55| = 0.07 (low)
  │  └─ final_ai_prob = 0.48 (use meta)
  │
  ├─ Forensic Calibration:
  │  ├─ No modules ≥ 0.60 (no strong support)
  │  ├─ Rule 3: final_ai_prob = 0.60*resnet + 0.40*weighted
  │  ├─ final_ai_prob = 0.60*0.32 + 0.40*0.43 = 0.364
  │
  ├─ Dynamic Threshold:
  │  ├─ No forensic support
  │  ├─ Model disagreement: 0.46 (moderate, triggers +0.10)
  │  ├─ Base threshold 0.68 + 0.10 = 0.78
  │
  ├─ Final Decision:
  │  ├─ ai_prob (0.364) < threshold (0.78)
  │  └─ DECISION: Real Image
  │  └─ (ViT overruled by ResNet + forensics agreement)
  │
  └─ Confidence: 65% real
     └─ "Model disagreement detected but ResNet
         confidence and weak forensics support
         real image classification"

Final Verdict: "Real Image"
  │ Confidence: 65%
  └─ Guard system prevented ViT false positive
\end{verbatim}

% ===============================================
\appendix
% ===============================================

\chapter{Appendix: Detailed Mathematical Formulas}
\label{app:formulas}

\section{SynthID Signal Score}

Composite calculation combining multiple quality metrics:

\begin{equation}
\text{signal\_score} = w_c \cdot f_c + w_p \cdot f_p + w_m \cdot f_m + w_s \cdot f_s + w_k \cdot f_k
\end{equation}

Where:
\begin{itemize}
    \item $f_c$ = normalized raw confidence
    \item $f_p$ = phase match (0-1)
    \item $f_m$ = correlation margin (normalized)
    \item $f_s$ = structure validity (0-1)
    \item $f_k$ = multi-scale consistency (0-1)
    \item $w_*$ = learned weights
\end{itemize}

\section{Frequency AI Score}

\begin{equation}
\text{ai\_score} = 0.40 \cdot \text{peak\_score} + 0.25 \cdot \text{pair\_score} + 0.25 \cdot \text{grid\_score} + 0.10 \cdot \text{align\_score}
\end{equation}

\section{Weighted Forensic Score}

\begin{equation}
\text{final\_score} = 0.50 \cdot s_r + 0.20 \cdot s_f + 0.20 \cdot s_a + 0.10 \cdot s_e
\end{equation}

Where $s_r$, $s_f$, $s_a$, $s_e$ are ResNet, Frequency, Anatomy, Entropy scores respectively.

\section{Confidence Scaling}

\begin{equation}
\text{confidence} = \begin{cases}
50 + 50 \cdot \frac{\text{ai\_prob} - \text{threshold}}{1.0 - \text{threshold}} & \text{if AI} \\
50 + 50 \cdot \frac{\text{threshold} - \text{ai\_prob}}{\text{threshold}} & \text{if Real}
\end{cases}
\end{equation}

\chapter{Appendix: File Locations Reference}
\label{app:files}

\section{Backend Files}

\begin{itemize}
    \item \texttt{backend/app.py} - Main Flask application
    \item \texttt{backend/combine\_model.py} - AI ensemble predictor
    \item \texttt{backend/custom\_forensics.py} - Forensic analysis modules
    \item \texttt{backend/forensic.py} - Report generation
    \item \texttt{backend/video\_detect\_standalone.py} - Video detection
    \item \texttt{backend/src/c2pa\_checker.py} - C2PA metadata verification
    \item \texttt{backend/src/synthid/service.py} - SynthID service
    \item \texttt{backend/src/synthid/robust\_extractor.py} - Watermark extraction
\end{itemize}

\section{Frontend Files}

\begin{itemize}
    \item \texttt{frontend/src/App.jsx} - Main router
    \item \texttt{frontend/src/pages/Home.jsx} - Landing page
    \item \texttt{frontend/src/pages/Dashboard.jsx} - Image analysis UI
    \item \texttt{frontend/src/pages/Report.jsx} - Image report
    \item \texttt{frontend/src/pages/VideoDashboard.jsx} - Video analysis UI
    \item \texttt{frontend/src/pages/VideoReport.jsx} - Video report
\end{itemize}

\section{Model Files}

\begin{itemize}
    \item \texttt{model\_output/resnet50\_finetuned\_benchmark.pth}
    \item \texttt{model\_output/ai\_detector\_meta\_learner.joblib}
    \item \texttt{model\_output/polynomial\_transformer.joblib}
    \item \texttt{model\_output/synthid/robust\_codebook.pkl}
    \item \texttt{backend/checkpoints/efficientnet.onnx}
    \item \texttt{backend/checkpoints/model.pth}
\end{itemize}

\end{document}
